Parsing Tweets into Universal Dependencies

### 2017-11-27 paper.tex
N - worked on sec2

\yjcomment{I've add the source of our example in the figure caption.}
\section{Annotation}\label{sec:anno}
\yicomment{Maybe we should contrast UD and kong's annotation strategy during each stage, because UD has also annotation guideline for every stage.}
In this section, we first review {\sc Tweebank v1} of \citet{kong-EtAl:2014:EMNLP2014},
which is the largest twitter dependencies annotation in previous literals (\S\ref{sec:tweebank}).
Then we introduce the difference in our tokenization and part-of-speech (re)annotation with \citet{ICWSM101540} and 
\citet{gimpel-EtAl:2011:ACL-HLT2011} on which {\sc Tweebank v1} was built (\S\ref{sec:pos-anno}). \yicomment{add tok} 
Then, we move to our effort of adopting the
UD conventions to cover tweet specific constructions. 
At last, we describe our process of creating new tweet treebank and characterize
the difficulties in reaching consistent annotations.
\nascomment{need to introduce this section and explain the structure.}
\yjcomment{done.}

@@

The annotation effort we describe stands in contrast to previous work
by \newcite{kong-EtAl:2014:EMNLP2014}.  Their aim was the rapid
development of a dependency parser for tweets, and to that end they
contributed a new annotated corpus, \textsc{Tweebank}, consisting of
929 tweets.  Their annotations added dependencies to a portion of the data
annotated with POS tags by 
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT} after tokenization \citep{ICWSM101540} \yicomment{add tok}.  
They also contributed a system for parsing;
we defer the discussion of their parser to \S\ref{sec:parsing}.

\citet{kong-EtAl:2014:EMNLP2014}'s rapid, small-scale annotation effort was heavily constrained.  It was
carried out mostly by non-native speakers, in a very short amount of
time (a day).  Driven both by the style of the text they sought to annotate
and by exigency, some of their annotation conventions included:
\begin{itemize}
\item Allowing an annotator to exclude tokens from the dependency
  tree.  A clear criterion for exclusion was not given, but many
  tokens were excluded because they were deemed ``non-syntactic.''
\item Allowing an annotator to merge a multiword expression into a
  single node in the dependency tree, with no internal structure.
  Annotators were allowed to take the same step with noun phrases.
\item Allowing multiple roots, since a single tweet might contain more
  than one sentence.
\end{itemize}
These conventions were justified on the grounds of making the
annotation easier for non-experts, but they must be revisited in our
effort to apply UD to tweets.

\subsection{Tokenization}\label{sec:tok-anno}
Our tokenization strategy is in the middle of the strategy of \citet{ICWSM101540} and that of UD. 
We want to fit Twitter texts better in UD dependencies while preserving the original tweets as much as possible.

\citet{ICWSM101540} preserve most of non-informal tokens including 
the hashtags, at-mentions, emoticons \& emojis and unicode glyphs. 
They also treat contractions and acronyms as whole tokens and do not split them.
UD tokenization\footnote{http://universaldependencies.org/u/overview/tokenization.html\yicomment{is it ok? no paper reference but webpage reference}} , in order to serve better dependency annotation, treat each syntactic word as token.
They take more aggressive strategy, including splitting off clitics from contractions or multi-word expressions ({\it gonna}) and combine several orthographic tokens into a single syntactic word. \yicomment{does UD support normalization for acronyms? Did not see it from website, need to check. related to ``idc''}

We do not touch any non-informal token in tweets and leave them as in the original tweets. 
Then we mainly follow UD tokenization. We tokenize contractions and multi-word expressions into their syntactic components, but do not combine tokens that have been already split in the original tweets back anymore.
For example, words like {\it gonna} or {\it can't} will be tokenized to {\it gon na} and {\it ca n't}, and parts like {\it Y O} and {\it R E T W E E T} will be not connected back to single words\footnote{Such tokens only accounts for 0.067\%, and we use {\it goeswith} relation to resolve it.}.

\subsection{Part-of-speech Annotation}\label{sec:pos-anno}

Before turning to UD annotations, we (re)annotated the data with 
POS, for consistency with other UD efforts,
which adopt the universal POS tagset of \newcite{PETROV12.274}.
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}

In some cases, conflicts arose between the UD English 
treebank conventions \cite[UD\_English]{Marneffe2014UniversalSD}
\footnote{\url{https://github.com/UniversalDependencies/UD_English}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT}.  In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD English treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

Compared to coarse-grained POS tagging effort of
\newcite{gimpel-EtAl:2011:ACL-HLT2011}, our approach simplifies some
matters.  For example, if a token is not considered syntactic by UD
conventions, it gets an \emph{other} (\texttt{X}) tag (Gimpel et
al.~had more extensive conventions).  Other phenomena, like
abbreviations, are more complicated for us, as discussed above;
Gimpel et al.~used a single part of speech for such expressions.

@@

Another important difference is in tokenization.  UD calls for more
aggressive tokenization than that used by
\newcite{gimpel-EtAl:2011:ACL-HLT2011}, which followed the rule-based
system  introduced by 
\newcite{ICWSM101540}.  In particular, they opted out of tokenizing
contractions and possessives, introducing new parts of speech
instead.\footnote{These tags only account for 2.7\% of tokens,
  leading to concerns about data sparseness in tagging and all
  downstream analyses.}  For us, these tokens must be split, but
universal parts of speech can be applied.

@@

\subsection{Universal Dependencies Applied to Tweets}

@@

\nascomment{for each thing we talk about here, might be nice to
  quantify it in our annotated corpus}

@@

\citet{eisenstein:2013:NAACL-HLT} studied the necessity of normalization
in social media text and pointed that such normalization is problematic.
Our solution to the syntax of abbreviations follows the spirit of his argument. Because abbreviations which clearly carry
syntactic functions only constitute 0.06\% of the tokens in our
dataset, we believe that normalization is an unnecessarily complicated step.
\nascomment{maybe we owe citations to
  \cite{finin-EtAl:2010:MTURK,eisenstein:2013:NAACL-HLT} here?  maybe
  better papers to cite}

\paragraph{Non-syntactic tokens.}  
\begin{table}
\centering
\begin{tabular}{lrr}
\multicolumn{2}{r}{syntactic (\%)} & non-syntactic (\%)\\ \hline
emoticons & 0.25 & 0.95 \\
RT & 0.14 & 2.49\\
hashtag & 1.02 & 1.24 \\
URL & 0.67 & 2.38 \\
truncated words & 0.00 & 0.49 \\
 \hline
 total & 2.08 & 7.55 \\
\end{tabular}
\caption{Proportion of non-syntactic tokens in the annotation.}\label{tbl:non-synt-prop}
\end{table}

@@

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			@username \& its \& \#awesome \& u \& gonna \& $\heart$ \& it \& Chk \& out \& our \& cooool \& project \& on \& http:// \& + \& RT \& pls\\
			\tiny PROPN \& \tiny PRON \& \tiny ADJ \& \tiny PRON \& \tiny VERB \& \tiny VERB \& \tiny PRON \& \tiny VERB \& \tiny ADP \& \tiny PRON \& \tiny ADJ \& \tiny NOUN \& \tiny ADP \& \tiny X \& \tiny SYM \& \tiny VERB\& \tiny INTJ\\
		\end{deptext}
		\deproot{3}{root}
		\depedge[edge unit distance=1em]{3}{1}{vocative}
		\depedge[edge unit distance=1em]{3}{2}{nsubj}
		\depedge[edge unit distance=1em]{3}{5}{parataxis}
		\depedge[edge unit distance=1em]{5}{4}{nsubj}
		\depedge[edge unit distance=1em]{5}{6}{xcomp}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.7em]{3}{8}{parataxis}
		\depedge[edge unit distance=1em]{8}{9}{compound:prt}
		\depedge[edge unit distance=0.7em]{8}{12}{obj}
		\depedge[edge unit distance=1em]{12}{10}{nmod:poss}
		\depedge[edge unit distance=1em]{12}{11}{amod}
		\depedge[edge unit distance=0.6em]{8}{14}{obl}
		\depedge[edge unit distance=1em]{14}{13}{case}
		\depedge[edge unit distance=0.55em]{8}{16}{conj}
		\depedge[edge unit distance=1em]{16}{15}{cc}
		\depedge[edge unit distance=0.37em]{3}{17}{discourse}
	\end{dependency}
	\caption{An example tweet with informal but syntactic tokens.
		This is an artificial tweets inspired by several tweets.}\label{fig:informal-toks}
\end{figure*}

The major characteristic that distinguishes tweets from standard texts is that
there are large proportion of tokens that don't carry any syntactic function.
In our annotation, there are five types of non-syntactic tokens: 
sentiment emoticons, retweet markers, topical hashtags, referential URLs, and
truncated words \footnote{Tweets we analyze have 140 character limits. Although Twitter has now doubled the tweet limit to 280 characters, we believe this type of tokens will still remain.}. Figure \ref{fig:non-syn-toks} illustrates examples of
these non-syntactic tokens.  As discussed above, these are generally
tagged with the \emph{other} (\texttt{X}) part of speech, except
emoticons, which are tagged as \emph{symbol} (\texttt{SYM}).  In our annotation, 
7.5\% of all tokens are non-syntactic; detailed statistics can be found in Table \ref{tbl:non-synt-prop}\footnote{The statistics are obtained on 140 character limit tweets.}.

@@

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions:
\begin{itemize}
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.
\end{itemize}

@@

\newcite{kong-EtAl:2014:EMNLP2014} proposed an additional
preprocessing step, \emph{token selection}, in their annotation process.
They required the annotators to first select the non-syntactic
tokens and exclude them from the final dependencies annotation.
%  \nascomment{confusing.  were they removed or
%  just labeled as non-syntactic and then excluded from the tree?  I
%  think they were not actually removed, and were still available,
%  e.g., for defining features of the input for parsing}
In order to keep our annotation conventions in line with UD norms,
we include non-syntactic tokens in our annotation following the
convention above. Compared with \citet{kong-EtAl:2014:EMNLP2014},
we also gave a clear definition of non-syntactic tokens, which helped
us avoid confusion during annotation.

@@

\subsection{\sc Tweebank v2}\label{sec:anno-process}
\yicomment{I think it should move to another section, just as in Kong's paper.}
Following the guidelines mentioned above, we create a new annotated Twitter treebank, which we call {\sc Tweebank v2}.

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1} (840 unique tweets, 639/201 for training/test set), along with additional 210\yicomment{changed to 210} tweets sampled from \newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled from the Twitter stream from February 2016 to July 2016\footnote{Data downloaded from \url{https://archive.org/}}.
The latter data source consists of 147.4M English tweets after being filtered by {\it lang} attribute in the tweet JSON and {\it langid.py}\footnote{\url{https://github.com/saffsd/langid.py}} toolkit.

@@

\subsubsection{Annotation Process}
Our annotation process was conducted in three stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the UD
annotation conventions and our guidelines.
Both the guidelines and annotation
were further refined by the authors of this paper to increase
the coverage of our guideline and solve inconsistency between
different annotators during this exercise. In the second stage, a POS tagger and
parser were trained on the annotated data from the first stage (1,041 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data. \nascomment{this is
  a little weird:  don't you have to do this before you parse?}\yjcomment{we were actually correcting the auto-analyzed 2,500 tweets.} Finally, an extra
layer of word-level normalization was manually annotated for data
analysis purposes.

@@

We report the inter-annotator agreement between the annotators
in the second stage. The agreement on POS is 96.61\%, the unlabeled
dependency agreement is 88.75\% and the labeled dependency agreement
is 84.31\%. Further analysis shows the major disagreements on POS
involve entity names (30.57\%) and topical hashtags (18.11\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of an empty auxiliary verb, or a clause-modified
noun, both annotations are reasonable.

>>>>

\yjcomment{I've add the source of our example in the figure caption.}
\section{Annotation}\label{sec:anno}
\yicomment{Maybe we should contrast UD and kong's annotation strategy during each stage, because UD has also annotation guideline for every stage.}

We first review {\sc Tweebank v1} of \citet{kong-EtAl:2014:EMNLP2014},
which is the largest Twitter dependencies annotation
(\S\ref{sec:tweebank}).
Then we introduce the differences in our tokenization
(\S\ref{sec:tok-anno}) and part-of-speech (\S\ref{sec:pos-anno}) (re)annotation with \citet{ICWSM101540} and 
\citet{gimpel-EtAl:2011:ACL-HLT2011}, respectively, on which {\sc Tweebank v1} was built. 
We describe our effort of adapting the
UD conventions to cover tweet-specific constructions (\S\ref{sec:ud-tweet}). 
Finally, we present our process of creating a new tweet treebank, {\sc
  Tweebank v2}, and characterize
the difficulties in reaching consistent annotations (\S\ref{sec:anno-process}).

@@

The annotation effort we describe stands in contrast to previous work
by \newcite{kong-EtAl:2014:EMNLP2014}.  Their aim was the rapid
development of a dependency parser for tweets, and to that end they
contributed a new annotated corpus, \textsc{Tweebank}, consisting of
929 tweets.  Their annotations added dependencies to a portion of the data
annotated with POS tags by 
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT} after rule-based tokenization \citep{ICWSM101540}.
Kong et al.~also contributed a system for parsing;
we defer the discussion of their parser to \S\ref{sec:parsing}.

Kong et al.'s rapid, small-scale annotation effort was heavily constrained.  It was
carried out mostly by nonnative speakers, in a very short amount of
time (a day).  Driven both by the style of the text they sought to annotate
and by exigency, some of their annotation conventions included:
\begin{itemize}
\item Allowing an annotator to exclude tokens from the dependency
  tree.  A clear criterion for exclusion was not given, but many
  tokens were excluded because they were deemed ``non-syntactic.''
\item Allowing an annotator to merge a multiword expression into a
  single node in the dependency tree, with no internal structure.
  Annotators were allowed to take the same step with noun phrases.
\item Allowing multiple roots, since a single tweet might contain more
  than one sentence.
\end{itemize}
These conventions were justified on the grounds of making the
annotation easier for non-experts, but they must be revisited in our
effort to apply UD to tweets.

\subsection{Tokenization}\label{sec:tok-anno}
Our tokenization strategy lies between the strategy of
\citet{ICWSM101540} and that of UD.
There is a significant tradeoff  between preservation of original tweet content and respecting
the UD guidelines.

The regex-based tokenizer of \citet{ICWSM101540}---which was 
originally designed for an exploratory search interface called
TweetMotif, not for NLP---preserves most whitespace-delimited tokens, including 
hashtags, at-mentions, emoticons,  emojis, and unicode glyphs. 
They also treat contractions and acronyms as whole tokens and do not split them.
UD
tokenization,\footnote{\url{http://universaldependencies.org/u/overview/tokenization.html}}
in order to better serve dependency annotation, treats each syntactic word as token.
They therefore take a more aggressive strategy, including splitting
clitics from contractions or multiword expressions (e.g., {\it
  gonna}), and sometimes combining several orthographic tokens into a
single syntactic word. \yicomment{does UD support normalization for
  acronyms? Did not see it from website, need to check. related to
  ``idc''} \nascomment{unless we make a change, I think we need to say
  here that acronyms like \emph{idc} are left as a single token}

In our manual tokenization, we do not modify any non-informal \nascomment{is this defined?} token in tweets and leave them as in the original tweets. 
Then we mainly follow UD tokenization. We tokenize contractions and
multiword expressions into their syntactic components, but do not
combine tokens that appear as split in the original tweet.
For example, words like {\it gonna} or {\it can't} will be tokenized
to {\it gon na} and {\it ca n't}, and sequences like {\it Y O} and
{\it R E T W E E T} will be not combined back into single
words.\footnote{Such tokens only account for 0.067\%; we use the UD
  {\it goeswith} relation to resolve these cases in the syntax.}

\subsection{Part-of-Speech Annotation}\label{sec:pos-anno}

Before turning to UD annotations, we (re)annotated the data with 
POS, for consistency with other UD efforts,
which adopt the universal POS tagset of \newcite{PETROV12.274}.
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}
In some cases, conflicts arose between the UD English 
treebank conventions \cite[][UD\_English]{Marneffe2014UniversalSD}\footnote{\url{https://github.com/UniversalDependencies/UD_English}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011} and
\newcite{owoputi-EtAl:2013:NAACL-HLT}.  In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD\_English treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

Compared to the coarse-grained POS tagging effort of
\newcite{gimpel-EtAl:2011:ACL-HLT2011}, our approach simplifies some
matters.  For example, if a token is not considered syntactic by UD
conventions, it gets an \emph{other} (\texttt{X}) tag (Gimpel et
al.~had more extensive conventions).  Other phenomena, like
abbreviations, are more complicated for us, as discussed above;
Gimpel et al.~used a single part of speech for such expressions.

@@

Another important difference follows from the difference in
tokenization.  As discussed above, UD calls for more
aggressive tokenization than that used by
\newcite{gimpel-EtAl:2011:ACL-HLT2011}.  In particular, Gimpel et al.~opted out of tokenizing
contractions and possessives, introducing new parts of speech
instead.\footnote{These tags only account for 2.7\% of tokens,
  leading to concerns about data sparseness in tagging and all
  downstream analyses.}  For us, these tokens must be split, but
universal parts of speech can be applied.

@@

\subsection{Universal Dependencies Applied to Tweets} \label{sec:ud-tweet}

@@

@@

\citet{eisenstein:2013:NAACL-HLT} studied the necessity of normalization
in social media text and argued that such normalization is problematic.
Our solution to the syntax of abbreviations follows the spirit of his argument. Because abbreviations which clearly carry
syntactic functions only constitute 0.06\% of the tokens in our
dataset, we believe that normalization is an unnecessarily complicated step.
% \nascomment{maybe we owe citations to
%   \cite{finin-EtAl:2010:MTURK,eisenstein:2013:NAACL-HLT} here?  maybe
%   better papers to cite}

\paragraph{Non-syntactic tokens.}  
\begin{table}
\centering
\begin{tabular}{lrr}
\multicolumn{2}{r}{syntactic (\%)} & non-syntactic (\%)\\ \hline
emoticons & 0.25 & 0.95 \\
RT & 0.14 & 2.49\\
hashtag & 1.02 & 1.24 \\
URL & 0.67 & 2.38 \\
truncated words & 0.00 & 0.49 \\
 \hline
 total & 2.08 & 7.55 \\
\end{tabular}
\caption{Proportions of non-syntactic tokens in our annotation.
These  statistics are obtained on 140 character-limited tweets. \label{tbl:non-synt-prop}}
\end{table}

@@

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			@username \& its \& \#awesome \& u \& gonna \& $\heart$ \& it \& Chk \& out \& our \& cooool \& project \& on \& http:// \& + \& RT \& pls\\
			\tiny PROPN \& \tiny PRON \& \tiny ADJ \& \tiny PRON \& \tiny VERB \& \tiny VERB \& \tiny PRON \& \tiny VERB \& \tiny ADP \& \tiny PRON \& \tiny ADJ \& \tiny NOUN \& \tiny ADP \& \tiny X \& \tiny SYM \& \tiny VERB\& \tiny INTJ\\
		\end{deptext}
		\deproot{3}{root}
		\depedge[edge unit distance=1em]{3}{1}{vocative}
		\depedge[edge unit distance=1em]{3}{2}{nsubj}
		\depedge[edge unit distance=1em]{3}{5}{parataxis}
		\depedge[edge unit distance=1em]{5}{4}{nsubj}
		\depedge[edge unit distance=1em]{5}{6}{xcomp}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.7em]{3}{8}{parataxis}
		\depedge[edge unit distance=1em]{8}{9}{compound:prt}
		\depedge[edge unit distance=0.7em]{8}{12}{obj}
		\depedge[edge unit distance=1em]{12}{10}{nmod:poss}
		\depedge[edge unit distance=1em]{12}{11}{amod}
		\depedge[edge unit distance=0.6em]{8}{14}{obl}
		\depedge[edge unit distance=1em]{14}{13}{case}
		\depedge[edge unit distance=0.55em]{8}{16}{conj}
		\depedge[edge unit distance=1em]{16}{15}{cc}
		\depedge[edge unit distance=0.37em]{3}{17}{discourse}
	\end{dependency}
	\caption{An example tweet with informal but syntactic tokens.
		This is a contrived example inspired by several tweets.}\label{fig:informal-toks}
\end{figure*}

The major characteristic that distinguishes tweets from standard texts is that
a large proportion of tokens don't carry any syntactic function.
In our annotation, there are five types of non-syntactic tokens: 
sentiment emoticons, retweet markers, topical hashtags, referential URLs, and
truncated words.\footnote{The tweets we analyze have at most 140
  characters. Although Twitter has doubled the tweet length limit to
  280 characters since our analysis, we believe this type of token
  will still remain.}  Figure \ref{fig:non-syn-toks} illustrates examples of
these non-syntactic tokens.  As discussed above, these are generally
tagged with the \emph{other} (\texttt{X}) part of speech, except
emoticons, which are tagged as \emph{symbol} (\texttt{SYM}).  In our annotation, 
7.5\% of all tokens are non-syntactic; detailed statistics can be found in Table \ref{tbl:non-synt-prop}.

@@

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions:
\begin{itemize}
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.
\end{itemize}

@@

\newcite{kong-EtAl:2014:EMNLP2014} proposed an additional
preprocessing step, \emph{token selection}, in their annotation process.
They required the annotators to first select the non-syntactic
tokens and exclude them from the final dependencies annotation.
%  \nascomment{confusing.  were they removed or
%  just labeled as non-syntactic and then excluded from the tree?  I
%  think they were not actually removed, and were still available,
%  e.g., for defining features of the input for parsing}
In order to keep our annotation conventions in line with UD norms,
we include non-syntactic tokens in our annotation following the
conventions above. Compared with \citet{kong-EtAl:2014:EMNLP2014},
we also gave a clear definition of non-syntactic tokens, which helped
us avoid confusion during annotation.

@@

\subsection{\sc Tweebank v2}\label{sec:anno-process}
Following the guidelines presented above, we create a new annotated Twitter treebank, which we call {\sc Tweebank v2}.

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1}
(840 unique tweets, 639/201 for training/test set), along with an 
additional 210 tweets sampled from the POS-tagged dataset of \newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled from the Twitter stream from February 2016 to July 2016.\footnote{Data downloaded from \url{https://archive.org/}.}
The latter data source consists of 147.4M English tweets after being
filtered by the {\it lang} attribute in the tweet JSON and {\it
  langid.py} toolkit.\footnote{\url{https://github.com/saffsd/langid.py}}

@@

\subsubsection{Annotation Process}
Our annotation process was conducted in three stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the UD
annotation conventions and our guidelines.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a POS tagger and
parser were trained on the annotated data from the first stage (1,041 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data. Finally, an extra
layer of word-level normalization was manually annotated for data
analysis purposes.

@@

We report the inter-annotator agreement between the annotators
in the second stage. The agreement on POS is 96.61\%, the unlabeled
dependency agreement is 88.75\% and the labeled dependency agreement
is 84.31\%. Further analysis shows the major disagreements on POS
involve entity names (30.57\%) and topical hashtags (18.11\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of an empty auxiliary verb, or a clause-modified
noun, either annotation is plausible.
